---
title: "Discard and bycatch monitoring program at industrial demersal fisheries in Chile: Where we are eight years from its implementation?"
#subtitle: A Short Subtitle
author:
  #     Another author footnote, this is a very long footnote and it should be a really long footnote. But this footnote is not yet sufficiently long enough to make two lines of footnote text.
  - name: Marcelo A. San Martin
    email: marcelo.sanmartin@ifop.cl
    attributes:
        corresponding: true
    note: This is the first author footnote.
    affiliations: 
        - id: some-tech
          name: Instituto de Fomento Pesquero (IFOP)
          department: Fisheries Department
          address: Blanco 839
          city: Valparaiso
          state: Chile
          postal-code: Chile
  - name: EL YEI CI
    affiliations: 
        - id: some-tech
  #- name: Juan Carlos Saavedra-Nievas
    #affiliations:
       # - id: some-tech
          # name: Instituto de Fomento Pesquero (IFOP)
          # department: Department Name
          # address: Street Address
          # city: City
          # state: State
          # postal-code: Postal Code
  #   note: |
  # - name: Cat Memes
  #   email: cat@example.com
  #   affiliations:
  #       - ref: another-u
  #   note: Yet another author footnote.
  # - name: Derek Zoolander
  #   email: derek@example.com
  #   affilations:
  #       - ref: some-tech
abstract: |
  Discard and bycatch have been a problem in world fisheries. During 2012, modifications on the Chilean Fisheries Law, including a permanent discard and bycatch research monitoring program (DBRMP) through scientific observers on-board in fisheries, were established to know and address this problem. With the aim of showing the steps and evolution of this process, indicators of discard and bycatch at demersal industrial fisheries between 2013 and 2020 years were assessed. Additionally, a summary of main regulatory measures were identified. The results showed that the involvement of fishermen was an important element to ensure the success of DBRMP and development of discard reduction plans. One of the biggest challenges to start the DBRMP was to achieve the engagement of them in a participatory process. Thus, meetings were held to provide information outreach about the law and DBRMP. After eight years from the beginning of the DBRMP, important improvements have been observed. Discard in trawling fisheries decreased on average 70% respect to initial values, while in long-line fisheries dropped 60%. Same trend was observed regarding bycatch of seabirds and sea lions. These results represent a key input to understand the discard problematic and the difficulties associated with their measurement, as well as the understanding of the underlying causes and solutions. Among the mandatory measures that has been implemented are; the discard ban of target species, adjustment of the non-target species catch rate, prohibition of bycatch, the mandatory use of bycatch mitigation measures systems as the grid device and tori lines used to avoid sea lions and seabirds, and the recent implementation of electronic monitoring. This work shows a brief overview of the main advances and solutions taken by Chilean fisheries government institutions regarding discard and bycatch, hoping to serve as a guideline for other countries which are advocated to address this matter.
  
keywords: 
  - Discard, bycatch
  - demersal fisheries 
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
editor: 
  markdown: 
    wrap: 72
---

**Introduction2**

Discard and bycatch have been a problem in world fisheries. Considering
the significant decreases in fish landing as of 1995 and increasing of
stocks overfished, during 2012, modifications on the Chilean Fisheries
Law with an ecosystem approach, including a permanent discard and
bycatch research monitoring program (DBRMP) through scientific observers
on-board in fisheries, were established to know and address this problem
(SanMartin2016; Roman2021). The main objectives of this monitoring
program included the determination of levels of discards in each
fishery, quantify bycatch (marine mammals, seabirds and marine turtles)
and identify the causes. Results obtained by DBRMP have been key to
implement the mandatory reduction measures to reduce the levels of
discard and bycatch at industrial demersal fisheries in Chile.

**Monitoring approach**

With the aim of showing the steps and evolution of this process,
indicators of discard and bycatch at demersal industrial fisheries
between 2013 and 2020 years were assessed. A total of ten industrial
fisheries, distributed from 28°S to 57°S, including longline and
trawling gear were considered (@fig-1).  Additionally, a summary of
main regulatory measures were identified.

![Chilean industrial demersal fisheries studied and spatial
coverage](fig1.png){#fig-1}



**Results**

The results showed that the involvement of fishermen was an important
element to ensure the success of DBRMP and development of discard
reduction plans. One of the biggest challenges to start the DBRMP was to
achieve the engagement of them in a participatory process. Thus,
meetings were held to provide information outreach about the law and
DBRMP. After eight years from the beginning of the DBRMP, important
improvements have been observed. Discard in trawling fisheries showed
variations in period evaluated, but in general decreased on average 70%
respect to initial values. The same trend was observed in long-line
fisheries, dropping around 60% with respect to the first years (Figure 2
A). Similar trend was observed regarding bycatch of seabirds and sea
lions, nonetheless, longline fisheries have not registered bycatch of
marine mammals and bycatch of seabird has been almost absent at
crustacean fisheries (Figure 2, B y C). Four general kinds of causes of
discard were identified; regulations, operational, quality and factors
associated with commercial issues. The last cause was the most important
with factors such as catch of non-commercial species and non-commercial
size. Bycatch of sea lions and seabirds were associated with
entanglement and when the animals feeding the catch or bait.

```{r}
#| warning: false
#| echo: false
library(readxl)
library(grid)
library(ggplot2)
```

```{r, fig.width=unit(10, "cm"), fig.height=unit(10, "cm")}
#| label: fig-hist-log-sa
#| out-width: 40%
#| fig-cap: "Discards."
#| warning: false
#| echo: false

desc<-read_excel("Data.xlsx", sheet ="Desc")
aves<-read_excel("Data.xlsx", sheet ="Aves_rate")
mam<-read_excel("Data.xlsx", sheet ="Mam_rate")

library(tidyverse)
library(reshape2)
library(dplyr)
desc.long<-melt(desc,id.vars=c(1),measure.vars=c(2:9),variable.name = "Anno",
          value.name = "Proporcion")#cap_total es la cpue estandarizada

pd <- position_dodge(0.5)
graf2<-ggplot(desc.long, aes(x=as.factor(Anno), y=Proporcion, group=Fleet, colour=Fleet)) +      
  geom_point(size=3,position=pd) + geom_line()+
  #geom_point(size=3,position=pd) + geom_line(arrow = arrow(),lwd=1)+
  #geom_errorbar(aes(ymin=l.lim, ymax=u.lim),linewidth=0.5, width=.1,position=pd)+
  labs(x="Año",
       y="Proporción descarte (%)",
       #title= "Estimación Total  crustáceos",
       subtitle="Descarte total",
       caption="Programa de Investigación Descarte Demersal",
       tag = "T: Trawling; L: Longline") +theme_bw()+
  scale_x_discrete(breaks =seq(2013,2020,1))+
  scale_y_continuous(limit = c(0,40),breaks=seq(0,40,10))+
  #facet_grid(~Proporción)
  facet_wrap(~Fleet,ncol=2)

win.graph(width=8000, height=6000,pointsize=10)
graf2+theme(legend.position = "none", plot.tag.position = c(0.7, 0.2), plot.margin=unit(c(1,1,1,1), "cm"))
```

# Bibliography styles

Here are two sample references: @Feynman1963118 @Dirac1953888
@SanMartin2016 @Roman2021.

By default, natbib will be used with the `authoryear` style, set in
`classoption` variable in YAML. You can sets extra options with
`natbiboptions` variable in YAML header. Example

```         
natbiboptions: longnamesfirst,angle,semicolon
```

There are various more specific bibliography styles available at
<https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files>.
To use one of these, add it in the header using, for example,
`biblio-style: model1-num-names`.

## Using CSL

If `cite-method` is set to `citeproc` in `elsevier_article()`, then
pandoc is used for citations instead of `natbib`. In this case, the
`csl` option is used to format the references. By default, this template
will provide an appropriate style, but alternative `csl` files are
available from <https://www.zotero.org/styles?q=elsevier>. These can be
downloaded and stored locally, or the url can be used as in the example
header.

# Equations

Here is an equation: $$ 
  f_{X}(x) = \left(\frac{\alpha}{\beta}\right)
  \left(\frac{x}{\beta}\right)^{\alpha-1}
  e^{-\left(\frac{x}{\beta}\right)^{\alpha}}; 
  \alpha,\beta,x > 0 .
$$

Inline equations work as well: $\sum_{i = 2}^\infty\{\alpha_i^\beta\}$

# Figures and tables

@fig-meaningless is generated using an R chunk.

```{r}
#| label: fig-meaningless
#| fig-cap: A meaningless scatterplot
#| fig-width: 5
#| fig-height: 5
#| fig-align: center
#| out-width: 50%
#| echo: false
plot(runif(25), runif(25))
```

# Tables coming from R

Tables can also be generated using R chunks, as shown in @tbl-simple
example.

```{r}
#| label: tbl-simple
#| tbl-cap: Caption centered above table
#| echo: true
knitr::kable(head(mtcars)[,1:4])
```

# References {.unnumbered}
